Probabilistic Inference of NYC Congestion from Taxi Data

### 2017-12-10 reports/paper/paper.tex
QED

Our inference strategy is based on a sampling-based approximating to the expectation over $\bm Z$. The key trick we use is that 
\begin{equation}
\label{eq:trick}
  \E_{Z}\bk{f(Z\uppr{k})} = |\bm Z\uppr{k}| \E_{\tilde Z}\bk{f(\tilde Z\uppr{k})p_Z(\tilde Z\uppr{k}|\bm W)},
\end{equation}
where $\tilde Z$ is drawn uniformly from $\bm Z$. This is the same technique as in importance sampling, except here the objective is to derive computationally tractable approximations, rather than maximizing the efficiency of the approximations. We note that the same trick allows us to replace any sum over $\bm Z$ with a weighted sum over the uniformly chosen sample. Exchanging $\log$ and expectation operator and applying \eqref{eq:trick} yields a lower bound for log-likelihood, up to a constant \[
-\frac{1}{L}\bk{\sum_{\tilde Z} \frac{1}{2\sigma^2} \pr{T - \sum_{i\in \tilde Z} w_i}^2 + \log \frac{\exp\pr{-\sum_i w_i}}{\sum_{\bm Z}\exp\pr{-\sum_i w_i}}}.
\]
Using $\tilde Z$ again to approximate the sum in the denominator yields the following upper bound for negative log-likelihood, up to a constant,
\begin{align}
&\frac{1}{2L\sigma^2}\sum_{j=1}^L \bk{T\uppr{k} + \sum_{i\in \tilde Z\uppr{k}_j}w_i}^2 + \frac{1}{L}\sum_{i=1}^L\sum_{i\in\tilde Z_{j}\uppr{k}}w_i 
\label{eq:softmax_objective}
\\&\qquad + \underset{i=1,\ldots,L}{\text{logsumexp}}\pr{-\sum_{\tilde Z_i} w_i} \nonumber,
\end{align}
where we sample uniformly and independently $\{\tilde Z_j\uppr{k}\}_{j=1}^L$. The pseudocode of the implementation is detailed in Algorithm~\ref{alg:main_alg}.

@@

@@

% table
%

Table~\ref{tab:mainres} is encouraging. Our models successfully leverage the
power of geography and pick up more signal in the data than distance of the trip
alone, as demonstrated by the significantly improved performance over linear
regression. Perhaps more strikingly, our model slightly outperforms the
predictions generated by Google Maps, which means that the predictive accuracy
of the model is on par with the standard of a widely used consumer product. Even
though we were not able to outperform the neural network, we reiterate that the
main advantage of our approach is to have interpretable parameters, which are
visualized in Figure~\ref{fig:unifmap}. We turn to the figure now.
% \begin{itemize}
% \item What conclusions can you draw from the results section?
% \item Is there further analysis you can do into the results of the system? Here is a good place to include visualizations, graphs, qualitative analysis of your results.
% \item  What questions remain open? What did you think might work, but did not?
% \end{itemize}

\begin{figure}[h!]
\centering
\includegraphics[width=0.5\textwidth]{figs/unifmaps.pdf}
\includegraphics[width=0.5\textwidth]{figs/softmaxmaps.pdf}
\caption{Learned weights, various times Tuesday. Top row: uniform model.
Bottom row: softmax model. \emph{In the grids, the southern border of Central Park lies
approximately three-sevenths of the way up}}
\label{fig:unifmap}
\end{figure}

@@

Finally, we note that our models condition only on hour and day of the week when
making predictions. However, our framework can incorporate a wider range of
relevant information. For instance, if weather data were accessible, the effect
of different weather conditions on road congestion could be measured. Similarly,
abnormal traffic conditions on holidays could be identified.

>>>>

Our inference strategy is based on a sampling-based approximating to the expectation over $\bm Z$. The key trick we use is that 
\begin{equation}
\label{eq:trick}
  \E_{Z}\bk{f(Z\uppr{k})} = |\bm Z\uppr{k}| \E_{\tilde Z}\bk{f(\tilde Z\uppr{k})p_Z(\tilde Z\uppr{k}|\bm W)},
\end{equation}
where $\tilde Z$ is drawn uniformly from $\bm Z$. This is the same technique as in importance sampling \cite{gelman2014bayesian}, except here the objective is to derive computationally tractable approximations, rather than maximizing the efficiency of the approximations. We note that the same trick allows us to replace any sum over $\bm Z$ with a weighted sum over the uniformly chosen sample. Exchanging $\log$ and expectation operator and applying \eqref{eq:trick} yields a lower bound for log-likelihood, up to a constant \[
-\frac{1}{L}\bk{\sum_{\tilde Z} \frac{1}{2\sigma^2} \pr{T - \sum_{i\in \tilde Z} w_i}^2 + \log \frac{\exp\pr{-\sum_i w_i}}{\sum_{\bm Z}\exp\pr{-\sum_i w_i}}}.
\]
Using $\tilde Z$ again to approximate the sum in the denominator yields the following upper bound for negative log-likelihood, up to a constant,
\begin{align}
&\frac{1}{2L\sigma^2}\sum_{j=1}^L \bk{T\uppr{k} + \sum_{i\in \tilde Z\uppr{k}_j}w_i}^2 + \frac{1}{L}\sum_{i=1}^L\sum_{i\in\tilde Z_{j}\uppr{k}}w_i 
\label{eq:softmax_objective}
\\&\qquad + \underset{i=1,\ldots,L}{\text{logsumexp}}\pr{-\sum_{\tilde Z_i} w_i} \nonumber,
\end{align}
where we sample uniformly and independently $\{\tilde Z_j\uppr{k}\}_{j=1}^L$. The pseudocode of the implementation is detailed in Algorithm~\ref{alg:main_alg}.

@@

\begin{figure}[h!]
\centering
\includegraphics[width=0.5\textwidth]{figs/unifmaps.pdf}
\includegraphics[width=0.5\textwidth]{figs/softmaxmaps.pdf}
\caption{Learned weights, various times Tuesday. Top row: uniform model.
Bottom row: softmax model.}
\label{fig:unifmap}
\end{figure}

@@

Table~\ref{tab:mainres} is encouraging. Our models successfully leverage the
power of geography and pick up more signal in the data than distance of the trip
alone, as demonstrated by the significantly improved performance over linear
regression. Perhaps more strikingly, our model slightly outperforms the
predictions generated by Google Maps, which means that the predictive accuracy
of the model is on par with the standard of a widely used consumer product. Even
though we were not able to outperform the neural network, we reiterate that the
main advantage of our approach is to have interpretable parameters, which are
visualized in Figure~\ref{fig:unifmap}.\footnote{In Figure~\ref{fig:unifmap}, the southern border of Central Park lies
approximately three-sevenths of the way up} We turn to the figure now.
% \begin{itemize}
% \item What conclusions can you draw from the results section?
% \item Is there further analysis you can do into the results of the system? Here is a good place to include visualizations, graphs, qualitative analysis of your results.
% \item  What questions remain open? What did you think might work, but did not?
% \end{itemize}

@@

Finally, we note that our models condition only on hour and day of the week when
making predictions. However, our framework can incorporate a wider range of
relevant information. For instance, if weather data were accessible, the effect
of different weather conditions on road congestion could be measured by adding the weather covariates to the Normal mean. Similarly,
abnormal traffic conditions on holidays could be identified.
